---
title: "Covid91 vaccine study Final2024"
author: "Sophia Castellon!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

## Methods 

Question : Is the COVID-91 vaccine effective in males?

$H_0$ The vaccine is not effective in preventing COVID-91 in males  

$H_A$ The vaccine is effective in preventing COVID-91 in males    

infected~sex(M)     
categorical~categorical

We will be using Cat~Cat
    
```{r}
Male <- FinalDat[FinalDat$sex=="M",]
```

```{r}
barchartGC(~infected + treatment,data=Male)
barchartGC(~infected + treatment,data=Male, type="percent")

```

The bar chart shows that you are less likely to contract COVID-91 if you did receive the vaccine. 

## Numerical Results 

```{r}
table1 <- xtabs(~infected + treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows the result in percentages. Within the placebo group 73.80% of individuals got infected with COVID-91. While in the vaccine group 26.20% contracted COVID-91. This indicates that receiving the vaccine as a male significantly reduces the risk of contracted COVID-91. 

## Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

This data and the fisher exact test shows that the P-Value is 2.2e-16 which is significantly lower than 0.05 which means we reject the null hypothesis. 

## Conclusion 

The bar chart shows that you are less likely to contract COVID-91 if you did receive the vaccine. The numerical data shows the results in percentages. Within the placebo group 73.80% of individuals got infected with COVID-91. This indicates that receiving the vaccine as a male significantly reduces the risk of contracted COVID-91. The inferential data including the Fisher exact test shows that the P-Value is 2.2e-16 which is significantly lower than 0.05 which means we reject the null hypothesis. All in all these examples and simulations show that as a male you are far less likely to contract COVID-91 upon receiving the vaccine. 

## Females

$H_0$ The vaccine is not effective in preventing COVID-91 in females  
$H_A$ The vaccine is effective in preventing COVID-91 in females   

```{r}
Female <- FinalDat[FinalDat$sex=="F",]
```

```{r}
barchartGC(~infected + treatment,data=Female)
barchartGC(~infected + treatment,data=Female, type="percent")

```



## LGBTQ

$H_0$ The vaccine is not effective in preventing COVID-91 in LGBTQ  
$H_A$ The vaccine is effective in preventing COVID-91 in LGBTQ   

## Druggies

$H_0$ The vaccine is not effective in preventing COVID-91 in druggies  
$H_A$ The vaccine is effective in preventing COVID-91 in druggies   

# Overall Results and Conclusions
