---
title: "Covid91 vaccine study Final2024"
author: "Sophia Castellon!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

## Methods 

Question : Is the COVID-91 vaccine effective in males?

$H_0$ The vaccine is not effective in preventing COVID-91 in males  

$H_A$ The vaccine is effective in preventing COVID-91 in males    

infected~sex(M)     
categorical~categorical

We will be using Cat~Cat
    
```{r}
Male <- FinalDat[FinalDat$sex=="M",]
```

```{r}
barchartGC(~infected + treatment,data=Male)
barchartGC(~infected + treatment,data=Male, type="percent")

```

The bar chart shows that you are less likely to contract COVID-91 if you did receive the vaccine. 

## Numerical Results 

```{r}
table1 <- xtabs(~infected + treatment, data=Male)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows the result in percentages. Within the placebo group 73.80% of individuals got infected with COVID-91. While in the vaccine group 26.20% contracted COVID-91. This indicates that receiving the vaccine as a male significantly reduces the risk of contracted COVID-91. 

## Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi square and p value is .00000000000000022 while the fisher exact test shows the p value is .00000000000000022 this is lower than .05 so we reject the null hypothesis. 

## Conclusion 

The bar chart shows that you are less likely to contract COVID-91 if you did receive the vaccine. The numerical data shows the results in percentages. Within the placebo group 73.80% of individuals got infected with COVID-91. This indicates that receiving the vaccine as a male significantly reduces the risk of contracted COVID-91. The inferential data including the Fisher exact test and chi square shows that the P-Value is .00000000000000022 which is significantly lower than 0.05 which means we reject the null hypothesis. All in all these examples and simulations show that as a male you are far less likely to contract COVID-91 upon receiving the vaccine. 

## Females

## Methods 

Question: Is the vaccine effective in preventing COVID-91 in feamles?

$H_0$ The vaccine is not effective in preventing COVID-91 in females  
$H_A$ The vaccine is effective in preventing COVID-91 in females  

infected~sex(F)   
categorical~categorical   

We will be using cat~cat

## Graphical results 

```{r}
Female <- FinalDat[FinalDat$sex=="F",]
```

```{r}
barchartGC(~infected + treatment,data=Female)
barchartGC(~infected + treatment,data=Female, type="percent")

```

This bar chart shows that females are less likely to contract COVID-91 after reviewing the vaccine.

## Numerical Results 

```{r}
table1 <- xtabs(~infected + treatment, data=Female)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows the results in percentages. Within the placebo group 64.93% of individuals got infected with COVID-91. While in the vaccinated group 35.07% contracted COVID-91. This indicates that receiving the vaccine as a female significantly reduces the risk of contracted COVID-91. 

## Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

This data and the Fisher exact test shows that the P-value is .00000001914 which is significantly lower than 0.05 which means we reject the null hypothesis. 

## Conclusion 

the bar chart shows that you are less likely to contract COVID-91 if you did receive the vaccine. The numerical data shows the results in percentages. Within the placebo group 64.39% of individuals got infected with COVID-91. While in the vaccine group 35.07% contracted COVID-91. This indicates that receiving the vaccine as a female significantly reduces the risk of contracted COVID-91. The inferential data including the Fisher exact test shows that the P-Value is 1.246e-08 which is significantly lower than 0.05 which means we reject the null hypothesis. In summation these examples and simulations show that is a female you are far less likely to contract COVID-91 if you receive the vaccine. 

## LGBTQ

## Methods 

$H_0$ The vaccine is not effective in preventing COVID-91 in LGBTQ

$H_A$ The vaccine is effective in preventing COVID-91 in LGBTQ  

## Graphical results 

```{r}
LGBTQ <- FinalDat[FinalDat$LGBTQ=="gay",]
```

```{r}
barchartGC(~infected + treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

The bar chart shows that members of the LGBTQ community are less likely to contract COVID-91 if you have received the vaccine. 

## Numerical Results 

```{r}
table1 <- xtabs(~infected + treatment, data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```
The numerical data shows that within the placebo group 80% of the LGBTQ community members got infected with COVID-91, while of the LGBTQ community members who received the vaccine only 20% got COVID-91. 

## Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi square test p value is .000002266 and the fisher exact test shows the p value is .0000007269. This is lower than 0.05 so we reject then null hypothesis. 

## Conclusion 

The bar chart shows that members of the LGBTQ community are less likely to contract COVID-91 if you received the vaccine. The numerical data shows that within the placebo group 80% of the LGBTQ community members got infected with COVID-91 while the LGBTQ community who received the vaccine only 20% got infected. The fisher exact test shows the p value is .0000007269 which is lower than 0.05 so we reject then null hypothesis.  

## Druggies

## Methods 

$H_0$ The vaccine is not effective in preventing COVID-91 in druggies  
$H_A$ The vaccine is effective in preventing COVID-91 in druggies 

## Graphical results 

```{r}
DrugUser <- FinalDat[FinalDat$DrugUser=="yes",]
```

```{r}
barchartGC(~infected + treatment,data=DrugUser)
barchartGC(~infected + treatment,data=DrugUser, type="percent")

```

The bar chart shows that members of the DrugUser community are more likely to contract COVID-91 if you have received the vaccine. 

## Numerical Results 

```{r}
table1 <- xtabs(~infected + treatment, data=DrugUser)
rowPerc(table1)
colPerc(table1)
```

The numerical data shows that within the placebo group 8.33% of the DrugUser community members got infected with COVID-91, while members who received the vaccine 91.67% got COVID-91. This indicates those who received the vaccine had a significant increase of contracting COVID-91. 

## Inferential Results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi square test shows the p value is .0000000001414 and the fisher exact test show the p value is .000000000003261 which is lower than .05 so we reject the null hypothesis. 

## Conclusion 

The bar chart shows that members of the DrugUsers community are more likely to contract COVID-91 if they received the vaccine. The numerical data shows that within the placebo group 8.33% of the DrugUser community members got infected, while the members who received the vaccine 91.76% got infected. This indicates those who recieved that vaccine had a significant increase of contracting COVID-91 The chi square test p value is .0000000001414 and the fisher exact test show the p value is .000000000003261 which is lower than .05 so we reject the null hypothesis. This data shows that DrugUsers are far more likely to contract COVID-91 if they received the vaccine. 

# Overall Results and Conclusions

Overall the data reflects that receiving the vaccine generally helps with preventing the contraction of COVID-91 if you are a female, male or LGBTQ. However, in the DrugUser community those who received the vaccine had a higher risk of contracting COVID-91 
